Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy

### 2024-02-04 Text/Articles/MIGARS_2024/Final_Version_MIGARS.Rmd
Corrected last details; now 4 pages

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

@@

## Statistical modeling of Intensity SAR data

@@

Simulations are conducted using $G_I^0$ distribution, with $200$ simulated samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\), with parameters \(\mu \in \left\{1, 10\right\}\), \(\alpha=-20\), and \(L=8\). 
In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 200

# Number of bootstrap replications
B <- 20
mu_values <- c(1, 10)
alpha <- -20
L <- 8

@@

calculate_results_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
  results_list <- list()

@@

    results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

@@

results_gi0 <- calculate_results_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=8$, $\\alpha=-20$.", fig.width=6, fig.height=5}

@@

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 2)

@@

print(
  kbl(
    reshaped_results,
    caption = "Bias and MSE of bootstrap estimators for $G_I^0$, $L=8$, $\\alpha=-20$.",
    format = "latex",
    booktabs = TRUE,
    align = "lrrrrrrrr",
    escape = FALSE, 
    digits = 3,  
    label = "table2"
  ) %>%
    add_header_above(c(" ", " ", "Bias" = 3, "MSE" = 3)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
    row_spec(0,  align = "c")
    #kable_styling(latex_options = "hold_position", "scale_down")
  )

@@

```{r Simulated_density, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
#, cache = TRUE, autodep = TRUE
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 1000
mu <- 1
B <- 2
alpha1 <- -8
sample.size <- c(50, 100, 200, 300)
L_values <- c(2, 5, 8, 18)

@@

load("./Data/results_5.Rdata")
summary_stats_5 <- all_summary_stats[["5"]]

@@

combined_summary_stats <- rbind(
  transform(summary_stats_2),
  transform(summary_stats_5),
  transform(summary_stats_8)

@@

table_result_combined <- kbl(
  combined_summary_stats,
  caption = "Descriptive analysis of $S(\\bm Z; L)$ for different values of $L$, and $\\mu=1$.",
  format = "latex",
  booktabs = TRUE,
  align = "lrrrrrrr",
  escape = FALSE,
  digits = 4,
  label = "table_stat_combined"
  #table.env = 'table*' # to span multiple columns
) %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

@@

After verifying the normality of the data, we proceeded to explore the capabilities of the proposed test in terms of size and power.
Under $\mathcal{H}_0$ the test statistic  is asymptotically distributed as $\mathcal{N}(0,\sigma_S^2))$, and therefore, the $p$-value of the observed test statistic is $2\Phi(\varepsilon)-1$, where $\Phi$ is the standard Gaussian cumulative distribution function and 
$$
\varepsilon=\frac{\tilde{H}(\bm{Z})-H_{\Gamma_{\text{SAR}}}(\bm{Z})}{\sigma_S}.
$$
We considered nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and \SI{10}{\percent}. 
In reference to the size, $1000$ simulations have been used for various sample sizes with  satisfactory results. In all cases, the nominal level has been achieved. 
The power, in general, improves with the increase in the sample size and the number of looks. 
The results are presented in Table&nbsp;\ref{tab:table_size_power}.

```{r Simulated_error_type_I, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

R <- 500
mu <- 1
B <- 2
L_values <- c(2, 5, 8)
sample_sizes <- c( 50, 100, 150, 200)
alpha_nominals <- c(0.01, 0.05, 0.1)

@@

```{r Simulated_power, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}

@@

        for (r in 1:R) {
          z <- gi0_sample(mu, -8, L, s)
          TestStat[r] <- bootstrap_correa_estimator_log_mean(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
        }

@@

R <- 500
mu <- 1
L_values <- c(2, 5, 8)
B <- 3
sample_sizes <- c(50, 100, 150, 200)
alpha_nominals <- c(0.01, 0.05, 0.1)

@@

R <- 500
mu <- 1
L <- 5
B <- 5
sample.size <- c( 81)

alpha1_values <- c( -8)

@@

```{r Plot_empirical_test, echo=FALSE, message=FALSE, warning=FALSE, out.width="90%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Empirical distributions under the null (centered around zero) and alternative hypotheses (e.g., $\\alpha=-8$, showing a shift from zero), with $\\mu=1$ and $L=5$."}

@@

for (alpha_val in alpha1_values) {
  plot <- ggplot(TestStatistics1, aes(x = Test_Statistics)) +
    geom_line(aes(col = factor(Sample_Size), linetype = factor(Sample_Size)), stat = "density", linewidth = 1.0) +
    geom_line(data = TestStatistics2, aes(x = Test_Statistics2), stat = "density", color = "blue", linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = difference.betweeen.GammaSAR.GI0.n.Look(-8, 5), col="darkgreen", linetype = "dotted", linewidth = 0.6) +
    annotate("text", x=difference.betweeen.GammaSAR.GI0.n.Look(-8, 5), y=6, parse=TRUE, label="alpha==-8", hjust=-.1, vjust=1.0, col="darkgreen") +
    geom_vline(xintercept = difference.betweeen.GammaSAR.GI0.n.Look(-1000, 5), col="darkgreen", linetype = "dotted", linewidth = 0.6) +
    theme_minimal() +
    theme(
      text = element_text(family = "serif"),
      legend.position = "none"  
    )+
    labs(x = "Test Statistics", y = "Density")

@@

```{r Real_SAR_Images, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center", fig.cap='Results of applying the test statistical to actual data.', fig.subcap=c('SAR image.', '$p$-values.' ), out.width='.48\\linewidth',  fig.ncol=2}
knitr::include_graphics("../../../Figures/PNG/Intensity.png")
knitr::include_graphics("../../../Figures/PNG/pvalue7x7.png")

>>>>

if(!require("rstudioapi")) install("rstudioapi")
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

@@

## Statistical Modeling of Intensity SAR Data

@@

Simulations are conducted using $G_I^0$ distribution, with $500$ simulated samples of size \(n\in\left\{9, 25, 49, 81, 121\right\}\). For brevity, we present the results with \(\mu=1\), \(\alpha=-10\), and \(L=5\). The results are consistent for other values of the mean. 
In the case of the bootstrap technique, each sample is replicated $100$ times with replacement.
We choose to use the following heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

@@

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121 )

# Number of replications
R <-500

# Number of bootstrap replications
B1 <- 100
mu_values <- c(1)
alpha <- -10
L <- 5

@@

calculate_results_gi0 <- function(sample_sizes, R, B1, mu_values, alpha, L, estimators) {
  results_list <- list()

@@

    results <- calculate_bias_mse_gi0(sample_sizes, R, B1, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

@@

results_gi0 <- calculate_results_gi0(sample_sizes, R, B1, mu_values, alpha, L, estimators)

@@

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=5$, $\\alpha=-10$.", fig.width=6, fig.height=3.0}

@@

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 1, nrow = 1)

@@

print(
  kbl(
    reshaped_results,
    caption = "Bias and MSE of bootstrap estimators for $G_I^0$, $L=5$, $\\alpha=-10$.",
    format = "latex",
    booktabs = TRUE,
    align = "lrrrrrrrr",
    escape = FALSE, 
    digits = 3,  
    label = "table2"
  ) %>%
    add_header_above(c(" ", " ", "Bias" = 3, "MSE" = 3)) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
    row_spec(0,  align = "c")
    #kable_styling(latex_options = "hold_position", "scale_down")
  )

@@

```{r Simulated_density, echo=FALSE, message=FALSE}
#, cache = TRUE, autodep = TRUE
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 800
mu <- 1
B <- 5

sample.size <- c(25,50, 100, 200, 300)
L_values <- c(2,  8)

@@

# load("./Data/results_5.Rdata")
# summary_stats_5 <- all_summary_stats[["5"]]

@@

combined_summary_stats <- rbind(
  transform(summary_stats_2),
  #transform(summary_stats_5),
  transform(summary_stats_8)

@@

table_result_combined <- kbl(
  combined_summary_stats,
  caption = "Descriptive analysis of $S(\\bm Z; L)$ with $L\\in\\left\\{2, 8\\right\\}$ and $\\mu=1$.",
  format = "latex",
  booktabs = TRUE,
  align = "lrrrrrrr",
  escape = FALSE,
  digits = 4,
  label = "table_stat_combined"
  #table.env = 'table*' # to span multiple columns
) %>%
  collapse_rows(latex_hline = "major", valign = "middle") %>%
  row_spec(0,  align = "c")

@@

After verifying the normality of the data, we proceeded to explore the capabilities of the proposed test in terms of size and power.
Under $\mathcal{H}_0$ the test statistic  is asymptotically distributed as $\mathcal{N}(0,\sigma_S^2)$, and therefore, the $p$-value of the observed test statistic is $2\Phi(\varepsilon)-1$, where $\Phi$ is the standard Gaussian cumulative distribution function and 
$$
\varepsilon=\frac{\tilde{H}(\bm{Z})-H_{\Gamma_{\text{SAR}}}(\bm{Z})}{\sigma_S}.
$$
We considered nominal levels of \SI{1}{\percent}, \SI{5}{\percent}, and \SI{10}{\percent}. 
In reference to the size, $1000$ simulations have been used for various sample sizes with  satisfactory results. In all cases, the nominal level has been achieved. 
The power, in general, improves with the increase in the sample size and the number of looks. 
The results are presented in Table&nbsp;\ref{tab:table_size_power}.

```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

R <- 1000
mu <- 1
B <- 2
L_values <- c( 5, 8)
sample_sizes <- c( 25, 50, 100, 200, 300)
alpha_nominals <- c(0.01, 0.05, 0.1)

@@

```{r Simulated_power, echo=FALSE, message=FALSE}

@@

        for (r in 1:R) {
          z <- gi0_sample(mu, -5, L, s)
          TestStat[r] <- bootstrap_correa_estimator_log_mean(z, B) - (L - log(L) + lgamma(L) + (1 - L) * digamma(L))
        }

@@

R <- 1000
mu <- 1
L_values <- c(5,  8)
B <- 1
sample_sizes <- c(25, 50, 100, 200, 300)
alpha_nominals <- c(0.01, 0.05, 0.1)

@@

R <- 2000
mu <- 1
L <- 8
B <- 1
sample.size <- c( 81)

alpha1_values <- c( -5)

@@

```{r Plot_empirical_test, echo=FALSE, message=FALSE, warning=FALSE, out.width="90%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Empirical distributions under the null (centered around zero) and alternative hypotheses (e.g., $\\alpha=-5$, showing a shift from zero), with $\\mu=1$ and $L=8$."}

@@

for (alpha_val in alpha1_values) {
  plot <- ggplot(TestStatistics1, aes(x = Test_Statistics)) +
    geom_line(aes(col = factor(Sample_Size), linetype = factor(Sample_Size)), stat = "density", linewidth = 1.0) +
    geom_line(data = TestStatistics2, aes(x = Test_Statistics2), stat = "density", color = "blue", linetype = "dashed", linewidth = 1) +
    geom_vline(xintercept = difference.betweeen.GammaSAR.GI0.n.Look(-5, 8), col="darkgreen", linetype = "dotted", linewidth = 0.6) +
    annotate("text", x=difference.betweeen.GammaSAR.GI0.n.Look(-5, 8), y=5, parse=TRUE, label="alpha==-5", hjust=-.1, vjust=1.0, col="darkgreen") +
    geom_vline(xintercept = difference.betweeen.GammaSAR.GI0.n.Look(-1000, 8), col="darkgreen", linetype = "dotted", linewidth = 0.6) +
    theme_minimal() +
    theme(
      text = element_text(family = "serif"),
      legend.position = "none"  
    )+
    labs(x = "Test Statistics", y = "Density")

@@

```{r Real_SAR_Images, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",fig.pos = 'H', fig.cap='Results of applying the test statistical to actual data.', fig.subcap=c('SAR image.', '$p$-values.' ), out.width='.48\\linewidth',  fig.ncol=2}
knitr::include_graphics("../../../Figures/PNG/Intensity.png")
knitr::include_graphics("../../../Figures/PNG/pvalue7x7.png")
